---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo = TRUE}
activity <- read.csv("activity.csv", header = TRUE, colClasses = c("numeric", "character", "numeric"), na = "NA")
```

## What is mean total number of steps taken per day?

# sum up steps by day

```{r, echo = TRUE}
StepDay <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

# plot histogram of total number of steps taken per day

```{r, echo = TRUE}
hist(StepDay, main = "Total Number of Steps Each Day", xlab = "Number of Steps", ylab = "Day")
```

## What is the average daily activity pattern?

# Mean and median of number of steps taken each day

```{r, echo = TRUE}
MeanDay <- mean(StepDay, na.rm = TRUE)
MedianDay <- median(StepDay, na.rm = TRUE)
```

# Time series plot of the average number of steps taken

```{r, echo = TRUE}
StepInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(as.numeric(names(StepInterval)), 
     StepInterval, 
     xlab = "5-Minute Interval", 
     ylab = "Steps", 
     main = "Average Number of Steps Daily", 
     type = "l")
```

# The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo = TRUE}
maxInterval <- names(sort(StepInterval, decreasing = TRUE)[1])
maxSteps <- sort(StepInterval, decreasing = TRUE)[1]
```

## Imputing missing values

# Code to describe and show a strategy for imputing missing data 

```{r, echo = TRUE}
StepInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```

# split activity data by interval

```{r, echo = TRUE}
activity.split <- split(activity, activity$interval)
```

# fill in missing data for each interval

```{r, echo = TRUE}
for(i in 1:length(activity.split)){
    activity.split[[i]]$steps[is.na(activity.split[[i]]$steps)] <- StepInterval[i]
}
activity.imputed <- do.call("rbind", activity.split)
activity.imputed <- activity.imputed[order(activity.imputed$date) ,]
```
# Histogram of the total number of steps taken each day after missing values are imputed

```{r, echo = TRUE}
StepDay.imputed <- tapply(activity.imputed$steps, activity.imputed$date, sum)
hist(StepDay.imputed, xlab = "Number of Steps", main = "Histogram: Steps per Day (Imputed data)")
```
## Are there differences in activity patterns between weekdays and weekends?

# Panel plot comparing the average number of steps taken per 5-minute interval across weekday days and weekends 

```{r, echo = TRUE}
activity.imputed$day <- ifelse(weekdays(as.Date(activity.imputed$date)) == "Saturday" | weekdays(as.Date(activity.imputed$date)) == "Sunday", "weekend", "weekday")
```

# Calculate average steps per interval for weekends

```{r, echo = TRUE}
StepInterval.weekend <- tapply(activity.imputed[activity.imputed$day == "weekend" ,]$steps, activity.imputed[activity.imputed$day == "weekend" ,]$interval, mean, na.rm = TRUE)
```

# Calculate average steps per interval for weekdays

```{r, echo = TRUE}
StepInterval.weekday <- tapply(activity.imputed[activity.imputed$day == "weekday" ,]$steps, activity.imputed[activity.imputed$day == "weekday" ,]$interval, mean, na.rm = TRUE)
```

# Set a 2 panel plot

```{r, echo = TRUE}
par(mfrow=c(1,2))
```

# Plot weekday activity

```{r, echo = TRUE}
plot(as.numeric(names(StepInterval.weekday)), 
     StepInterval.weekday, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Daily Activity – Weekday", 
     type = "l")
```

# Plot weekend activity

```{r, echo = TRUE}
plot(as.numeric(names(StepInterval.weekend)), 
     StepInterval.weekend, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Daily Activity – Weekend", 
     type = "l")
```
